---
title: "Bengaluru House Price Analytics Report"
author: 'Name : Sanjay Kumar Sinha          Roll Number : MDS202230'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

# Introduction

In this project we analyse the key things that a potential home buyer considers before purchasing a house. The location, the size of the property, vicinity to offices, schools and the most important factor - the price.

Buying a house has always been a tricky choice, especially in a metro city like Bengaluru. With its millennial crowd, vibrant culture, great climate and a slew of job opportunities, it is difficult to ascertain the price of a house.


# Dataset Description

The dataset has details of some of the houses in Bengaluru spread across different locations. It originally had 9 columns and 13320 rows. Some of the attributes are location of the house, area in sqft, number of bathroom, number of balcony, price, etc.

We cleaned the data and preprocessed it and are currently dealing with 8 variables out of which 5 (area_type, BHK, bath, balcony and location) are categorical and rest 3 (price, area_sqft, price_per_sqft) are numerical

The data has been obtained from the website 
https://www.kaggle.com/datasets/amitabhajoy/bengaluru-house-price-data



```{r, echo=FALSE}
# Importing libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(tm)
library(stringr)
library(corrplot)
library(tidyverse)


data = read.csv(file = "Bengaluru_House_Data.csv", header = T)
df = data.frame(data)
#head(df)
#summary(df)
theme_update(plot.title= element_text(hjust = 0.5, size =10))
```


```{r, echo=FALSE}
# Finding missing values and cleaning data frame
df <- replace(df, df=='', NA)

missing_values = colSums(is.na(df))# missing values in each column
#missing_values

df[,c(2,5)] = list(NULL)
df = na.omit(df)

```



```{r ,warning=FALSE, echo=FALSE}
# Pre-Processing the data

# create BHK from size column in numeric type

df$BHK <- 0
df$BHK = extract_numeric(df$size)  # using tidyr
df = df[,-3]


# total_sqft column in required format


df = subset(df, removeNumbers(data$total_sqft) == "")  
df$area_sqft = as.numeric(gsub(".*?([0-9]+).*", "\\1", df$total_sqft))
df = df[,-3]

df$price_per_sqft = df$price / df$area_sqft

```



```{r, echo=FALSE}
# Removing Outlyers

# delete rows where bath > bhk
df = subset(df, df$bath <= df$BHK)

# delete rows where bhk > 5
df = subset(df, df$BHK <= 5)

# delete rows where area_sqft > 4000
df = subset(df, df$area_sqft <= 4000)
```


```{r, echo=FALSE}
# Reconverting BHK, bath, balcony into catregorical form
df$BHK = as.character(df$BHK)
df$bath = as.character(df$bath)
df$balcony = as.character(df$balcony)

```






```{r, echo=FALSE }
# Including Plots

# Pie Chart of BHK

df1=df %>% group_by(BHK) %>% summarise(count=n()) 
df1$count = df1$count/sum(df1$count)*100
df1$count = round(df1$count, digits = 2)
df1$BHK = factor(df1$BHK,levels= as.character(1:5))
ggplot(df1, aes(x="",y=count, fill=BHK)) +  geom_col()+ geom_text(aes(x=1.56,label= count), size=3, angle = 60, position = position_stack(vjust=0.5))+coord_polar(theta='y', direction = -1)+theme_void() +ggtitle('Pie Chart of BHK')+ theme(plot.title = element_text(size=15))




```





Here we observe that most of the houses are either 2-BHK or 3-BHK



```{r, echo=FALSE, warning = FALSE, fig.width = 15,     fig.height=10}
# Scatterplot House Area v/s Price
ggplot(df, aes(x= area_sqft, y=price, color=area_type)) + geom_jitter() + ylim(0,500) +facet_wrap(~area_type) + 
labs(x='Area (sq. ft)', y='Price (in lacs)', title='Area v/s Price Scatterplot')+ theme_grey(base_size = 30)+ theme(legend.position = 'None')

```





From the scatterplot we observe that price of the house increases as the area of the house increaes


```{r, echo=FALSE, warning = FALSE, fig.width = 20, fig.height=15}
# Boxplot BHK v/s Price

ggplot(df, aes(x = BHK , y = price, 
                             fill = BHK)) +
    theme(legend.position = 'None')+
    xlab("BHK")+ ylab("Price (in lacs)") +
    geom_boxplot() + stat_boxplot(geom = 'errorbar')+
      ggtitle('BHK v/s Price Boxplot')+ ylim(0,500)+
  theme(plot.title = element_text(size=50), axis.text =  element_text(size=30), axis.title =  element_text(face='bold',size=40))

```




From the boxplot we observe that the price of 1-BHK, 2-BHK and 3-BHK houses does not vary much compared to 4-BHK and 5-BHK. Therefore we conclude that as the BHK increases, variation in price also increases. Moreover price of 5-BHK houses are less than that of 4-BHK. It is mainly due to the location factor.





```{r, echo=FALSE}

# Heatmap of variables
df2 = data.frame(df)
df2$BHK = as.numeric(df2$BHK)
df2$bath = as.numeric(df2$bath)
df2$balcony = as.numeric(df2$balcony)
num_cols = sapply(df2,is.numeric)
num_cols = num_cols[-6]
cor_data = cor(df2[,num_cols])
corrplot(cor_data, method = "color", title = 'Correlation Heatmap', mar=c(0,0,1,0))


```




From the heatmap we observe that the Area of the house is positively correlated wih its Price as we expect it to be. Also the number of bathrooms in a house is strongly positively correlated with BHK.
