---
title: "Bengaluru House Price Analytics Report"
author: 'Name : Sanjay Kumar Sinha   Roll Number : MDS202230'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

## In this project we analyse the key things that a potential home buyer considers before purchasing a house. The location, the size of the property, vicinity to offices, schools and the most important factor - the price.

## Buying a house has always been a tricky choice, especially in a metro city like Bengaluru. With its millennial crowd, vibrant culture, great climate and a slew of job opportunities, it is difficult to ascertain the price of a house in Bengaluru.


# Dataset Description

## The dataset originally had 9 columns and 13320 rows

## The data has been obtained from the website 
## https://www.kaggle.com/datasets/amitabhajoy/bengaluru-house-price-data



```{r, echo=FALSE}
# Importing libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(tm)
library(stringr)
library(corrplot)


data = read.csv(file = "Bengaluru_House_Data.csv", header = T)
df = data.frame(data)
head(df)
summary(df)
```

```{r, echo=FALSE}
# Finding missing values and cleaning data frame
df <- replace(df, df=='', NA)

missing_values = colSums(is.na(df))# missing values in each column
missing_values

df[,c(2,5)] = list(NULL)
df = na.omit(df)

```



```{r ,warning=FALSE, echo=FALSE}
# Pre-Processing the data

# create BHK from size column in numeric type

df$BHK <- 0
df$BHK = extract_numeric(df$size)  # using tidyr
df = df[,-3]


# total_sqft column in required format


df = subset(df, removeNumbers(data$total_sqft) == "")  
df$area_sqft = as.numeric(gsub(".*?([0-9]+).*", "\\1", df$total_sqft))
df = df[,-3]

```



```{r, echo=FALSE}
# Removing Outlyers

# delete rows where bath > bhk
df = subset(df, df$bath <= df$BHK)

# delete rows where bhk > 10
df = subset(df, df$BHK <= 10)

# delete rows where area_sqft > 15000
df = subset(df, df$area_sqft <= 15000)
```





```{r, echo=FALSE}
# Including Plots

```


